Seasonality of submesoscale dynamics in the Kuroshio Extension

### 2016-10-11 writeup/RochatEtal_ResponseToReviewers_BIS.tex

\bdp
   \item Changed.
\edp

\item {\it Line 187. I am not sure if PSI is important near surface and wonder if it
                is resolved in this model because of the vertical resolution.
                Authors could discuss about PSI in the model with Jenn Mackinnon there.}

\item {\it Line 196. Remove "?"}

\bdp
  \item A reference  was amiss --- the ``?'' was raised during the\, \LaTeX\, compilation. Fixed.
\edp

@@

      \bdp
        \item Thanks. Better to use APV.
      \edp

\end{enumerate}

\section{Reviewer \#3}

{\it The manuscript provides model-based evidence of a seasonal cycle in submesoscale motions
in the western North Pacific (Kuroshio Extension). The manuscript summarizes the results of
a study in which two years of moderate-resolution (1/24th degree or 4.6 km) and one year of
high-resolution (1/48th degree or 2.3 km) model output were examined for evidence of
submesoscale activity. Measures of submesoscale activity include vorticity, divergence and
strain rate and the authors demonstrate that these quantities are enhanced in winter. Kinetic
energy and sea surface height (SSH) spectra also show seasonal trends. These results are
consistent with an existing suite of studies reporting seasonality to submesoscale flows. What
is relevant for the present study and what distinguishes it from previous studies is that the
authors report an increase in internal wave activity during summer and which projects onto
the SSH variance. Because of the relevance of this study to challenges facing the upcoming
Surface Water Ocean Topography (SWOT) mission—which aims at measuring submesoscale
flows from space from nearly-balanced flows—this study will be of interest to readers of
GRL.

This is the second submission of the manuscript. The errors in the previous manuscript have
been addressed. The authors provide a direct connection between inertia-gravity waves and
SSH variance (Figure 4 of their manuscript) and illustrate that these undergo seasonality.
They have also framed the manuscript in terms of SWOT, which helps the reader understand
the importance of the results. There is no explicit connection between divergence and inertia gravity
waves (IGWs), making the connection between the two segments of the manuscript
tentative. (Part 1: analysis of the gradient tensor. Part 2: spectral analysis of SSH.) That is,
other ageostrophic flows can contribute to this divergence. Nevertheless, the seasonality in the
internal wave field is evident, complementing the observed seasonality in vorticity,
divergence and strain rate. In summary, despite the potential for varying internal wave
activity to be specifically tied to the Kuroshio Extension region, the implication of seasonal
variation in internal wave activity is new, and I recommend publication pending a few
revisions listed below.}

\bdp
   \item We thank the reviewer for the thorough and constructive review.
\edp

\begin{enumerate}
 \item {\it I appreciate the changes made by the authors. I am happy to know that the suggestions in
the previous review helped. Incidentally, I also appreciated the response by the authors
regarding the resolution of the model and scales energized by baroclinic instability, citing
work by Callies et al. (2016) and Larichev and Held (1995). Lastly, I am grateful for the
authors’ acknowledgement toward the end of the manuscript. However, I would rather the
authors remove me from the acknowledgements since I am involved in the formal review
process. There is no problem with this and I appreciate the authors’ efforts to thank me.}

\bdp
   \item We have removed the referee's name from the acknowledgements.
\edp

\bdp

\item {\it 2. I have been looking at the $\omega-k$ spectra in Figure 4. There is a lot of interesting
information in these data. In particular, I have found three characteristics of interest. See
attached figure for my definition of corresponding regions.}

    \subitem (a) {\it Region 1: If one interprets submesoscale but larger-period waves (e.g., $\omega$ < f32.5 \& $k$ >
10-2 km-1
) as associated with gradient wind- or geostrophically-balanced flows and/or
front-related flows (note: I am not certain of this interpretation1
), there is considerable
change in energy from April to October, reflective of frontogenetic activity being
enhanced in late winter/early spring. This is consistent with the authors’ depiction of
surface fields in Figure 1d and 1e.}

  \subitem (b) {\it Region 2: Another interesting feature is the lack of energy in the inertia-gravity wave
(IGW) band (i.e., between mode-1 at f40 and mode-4 at f25) in April but presence of
this energy in October. This is a nice result.}

\subitem (c) {\it Region 3: Finally, the last feature I note is the change in energy at $\omega$ = 2f32.5 and at
horizontal wavenumbers $k$ > 8 x10-3 km-1 (i.e., horizontal wavelengths < 125 km).
Although these are IGWs themselves and may result from increased stratification as
mentioned by the authors, there still does seem to be elevated energy at 2f during
October. As one does not expect internal waves at the M2 frequency to change with
season, this seasonal change in energy is most likely explained by wave-wave
interactions. I confess I don’t know much about this topic but there is the suggestion
of such waves generated as a result of near-inertial wave interaction with quasigeostrophic
flows (Wagner and Young 2016). Perhaps the authors could discuss this
with GW or BY since they are mentioned in the acknowledgements. Niwa and Hibiya
(1999) also document such motions generated by an atmospheric disturbance and
which Olbers (1983) suggests can propagate very long distances. (Is it possible that
typhoon-related depressions that cross over the Kuroshio during October but not in
April excite inertial motions of frequency 2f and that then propagate into the analysis
region? The authors need not answer this in their paper ... Food for thought.) In
summary, I think (1) it would be worth mentioning that the energy at $\omega$ = 2f is
seasonally varying, as the authors are documenting, (2) that this projects onto SSH
variance at small horizontal scales, and (3) will be a source of energy seen by the
SWOT altimeter. This may be a result specific to the Kuroshio region but is
nevertheless interesting.}

\edp

\bdp
\item {\it I recommend the authors use a different colormap when plotting spectra in Figure so that the
reader can see changes in internal wave energy more easily. Or perhaps contour levels?}
\edp

\bdp
\item {\it As a side note, Reviewer 2 made a comment in the previous draft about near-inertial waves
(NIWs). S/he hypothesized that they may be an important source of SSH variance. Given the
above conversation, this Reviewer may be at least partly correct. It seems, though, as you say
that the waves near f penetrate deep into the bottom, since there is virtually no SSH variance
here}
\edp

\item{\it In the previous submission, I suggested estimating the magnitude of the sea surface height
signature within a particular band as a function of season. You have done this and it is nice.
Table 1 is also helpful for placing these quantities into context.

2 ne question I have though is that, given that the bands described above have rather
complicated shape in $\omega-k$ space, might you instead integrate the PSD over these contoured
regions to determine the SSH variance found within each band and corresponding to each
type of motion? It would be interesting, for example, to know what fraction of the SSH signal
could be directly attributed to (1) IGWs motions within the dispersion relation curves and for
$k$ > 10-2 km-1 (region 2) (2) motions near $\omega$ = 2f32.5 \& $k$ > 10-2 km-1 (region 3) and (3) the
other portion of the spectra (region 1), which may reflect balanced and nearly-balanced
submesoscale motions. While the authors may not anticipate understanding energy within
these bands completely, future scientists may cite these motions as contributing to an
appreciable amount of the submesoscale SSH signal. If the authors feel this is not necessary
to include within the manuscript, this could be placed within the supporting material since it
would be interesting to the SWOT community}

\item{\it  There are a few wording mistakes that the authors should address prior to resubmission. I
don’t list them all but here are a few:}

  \subitem (a) {\it Figure 4, caption: “For reference, the gray curvish line represents” ==> Reword as,
  “For reference, we provide the baseline requirement as specified by the SWOT
  science team”, and provide a formal reference to where the reader can find this.}

  \subitem (b) {\it   Line 168: “The Similarly” ==> “Similarly...''}

  \subitem (c) {\it Lines 31-35: The revisit coverage of the SWOT spacecraft is on the order of 11 days
  (https://swot.jpl.nasa.gov/files/swot/SRD$\_$021215.pdf). I don’t really see how the
  inability to decouple geostrophic and ageostrophic motions presents an opportunity.
  But that being said, I would eliminate hyphens from line 33. It seems very informal.
  The colons in the sentence following this are fine.}

### 2016-10-11 writeup/RochatEtal_ResponseToReviewers_BIS.tex

\item {\it I recommend the authors use a different colormap when plotting spectra in Figure so that the
reader can see changes in internal wave energy more easily. Or perhaps contour levels?}

\item {\it As a side note, Reviewer 2 made a comment in the previous draft about near-inertial waves
(NIWs). S/he hypothesized that they may be an important source of SSH variance. Given the
above conversation, this Reviewer may be at least partly correct. It seems, though, as you say
that the waves near f penetrate deep into the bottom, since there is virtually no SSH variance
here}

\item{\it In the previous submission, I suggested estimating the magnitude of the sea surface height
signature within a particular band as a function of season. You have done this and it is nice.
Table 1 is also helpful for placing these quantities into context.

One question I have though is that, given that the bands described above have rather
complicated shape in $\omega-k$ space, might you instead integrate the PSD over these contoured
regions to determine the SSH variance found within each band and corresponding to each
type of motion? It would be interesting, for example, to know what fraction of the SSH signal
could be directly attributed to (1) IGWs motions within the dispersion relation curves and for
$k$ > 10-2 km-1 (region 2) (2) motions near $\omega$ = 2f32.5 \& $k$ > 10-2 km-1 (region 3) and (3) the
other portion of the spectra (region 1), which may reflect balanced and nearly-balanced
submesoscale motions. While the authors may not anticipate understanding energy within
these bands completely, future scientists may cite these motions as contributing to an
appreciable amount of the submesoscale SSH signal. If the authors feel this is not necessary
to include within the manuscript, this could be placed within the supporting material since it
would be interesting to the SWOT community}

### 2016-10-13 writeup/RochatEtal_ResponseToReviewers_BIS.tex

\begin{enumerate}
 \item {\it I appreciate the changes made by the authors. I am happy to know that the suggestions in
the previous review helped. Incidentally, I also appreciated the response by the authors
regarding the resolution of the model and scales energized by baroclinic instability, citing
work by Callies et al. (2016) and Larichev and Held (1995). Lastly, I am grateful for the
authors’ acknowledgement toward the end of the manuscript. However, I would rather the
authors remove me from the acknowledgements since I am involved in the formal review
process. There is no problem with this and I appreciate the authors’ efforts to thank me.}

\bdp
   \item We have removed the referee's name from the acknowledgements.
\edp

\item {\it 2. I have been looking at the $\omega-k$ spectra in Figure 4. There is a lot of interesting
information in these data. In particular, I have found three characteristics of interest. See
attached figure for my definition of corresponding regions.}

    \subitem (a) {\it Region 1: If one interprets submesoscale but larger-period waves (e.g., $\omega < f_{32.5}$ \& $k >
10-2$ km$^{-1}$
) as associated with gradient wind- or geostrophically-balanced flows and/or
front-related flows (note: I am not certain of this interpretation1
), there is considerable
change in energy from April to October, reflective of frontogenetic activity being
enhanced in late winter/early spring. This is consistent with the authors’ depiction of
surface fields in Figure 1d and 1e.}

  \subitem (b) {\it Region 2: Another interesting feature is the lack of energy in the inertia-gravity wave
(IGW) band (i.e., between mode-1 at $f_{40}$ and mode-4 at $f_{25}$) in April but presence of
this energy in October. This is a nice result.}

\subitem (c) {\it Region 3: Finally, the last feature I note is the change in energy at $\omega$ = 2f32.5 and at
horizontal wavenumbers $k > 8 \times 10-3$ km$^{-1}$ (i.e., horizontal wavelengths $< 125$ km).
Although these are IGWs themselves and may result from increased stratification as
mentioned by the authors, there still does seem to be elevated energy at $2f$ during
October. As one does not expect internal waves at the M2 frequency to change with
season, this seasonal change in energy is most likely explained by wave-wave
interactions. I confess I don’t know much about this topic but there is the suggestion
of such waves generated as a result of near-inertial wave interaction with quasigeostrophic
flows (Wagner and Young 2016). Perhaps the authors could discuss this
with GW or BY since they are mentioned in the acknowledgements. Niwa and Hibiya
(1999) also document such motions generated by an atmospheric disturbance and
which Olbers (1983) suggests can propagate very long distances. (Is it possible that
typhoon-related depressions that cross over the Kuroshio during October but not in
April excite inertial motions of frequency $2f$ and that then propagate into the analysis
region? The authors need not answer this in their paper ... Food for thought.) In
summary, I think (1) it would be worth mentioning that the energy at $\omega$ = $2f$ is
seasonally varying, as the authors are documenting, (2) that this projects onto SSH
variance at small horizontal scales, and (3) will be a source of energy seen by the
SWOT altimeter. This may be a result specific to the Kuroshio region but is
nevertheless interesting.}

@@

\item {\it I recommend the authors use a different colormap when plotting spectra in Figure so that the
reader can see changes in internal wave energy more easily. Or perhaps contour levels?}

\item {\it As a side note, Reviewer 2 made a comment in the previous draft about near-inertial waves
(NIWs). S/he hypothesized that they may be an important source of SSH variance. Given the
above conversation, this Reviewer may be at least partly correct. It seems, though, as you say
that the waves near f penetrate deep into the bottom, since there is virtually no SSH variance
here}

\item{\it In the previous submission, I suggested estimating the magnitude of the sea surface height
signature within a particular band as a function of season. You have done this and it is nice.
Table 1 is also helpful for placing these quantities into context.

One question I have though is that, given that the bands described above have rather
complicated shape in $\omega-k$ space, might you instead integrate the PSD over these contoured
regions to determine the SSH variance found within each band and corresponding to each
type of motion? It would be interesting, for example, to know what fraction of the SSH signal
could be directly attributed to (1) IGWs motions within the dispersion relation curves and for
$k > 10-2$ km$^{-1}$ (region 2) (2) motions near $\omega$ = 2$f_{32.5}$ \& $k$ > 10-2 km$^{-1}$ (region 3) and (3) the
other portion of the spectra (region 1), which may reflect balanced and nearly-balanced
submesoscale motions. While the authors may not anticipate understanding energy within
these bands completely, future scientists may cite these motions as contributing to an
appreciable amount of the submesoscale SSH signal. If the authors feel this is not necessary
to include within the manuscript, this could be placed within the supporting material since it
would be interesting to the SWOT community}

@@

  \subitem (b) {\it   Line 168: “The Similarly” ==> “Similarly...''}

  \bdp
      \item \noindent Fixed.
  \edp

  \subitem (c) {\it Lines 31-35: The revisit coverage of the SWOT spacecraft is on the order of 11 days
  (https://swot.jpl.nasa.gov/files/swot/SRD$\_$021215.pdf). I don’t really see how the
  inability to decouple geostrophic and ageostrophic motions presents an opportunity.
  But that being said, I would eliminate hyphens from line 33. It seems very informal.
  The colons in the sentence following this are fine.}

  \bdp
    \item {\noindent We removed the dashes. There are many opportunities associated with observations
          of ageostrophic submesoscale SSH from SWOT. At very least, the projection of both geostrophic
          and ageostrophic flows onto SSH --- and the interest in separating those flows --- will boost research
          on interactions between those types of motion. We have changed that sentence to make this point clear. }
  \edp

### 2016-10-13 writeup/rocha_etal.tex
Minor fix in caption
Small fixed

\begin{abstract}
    The vigorous seasonal cycle of the mixed layer modulates upper-ocean submesoscale turbulence.
    Here we provide model-based evidence that the seasonally-changing upper-ocean
    stratification in the Kuroshio Extension also modulates submesoscale  (here 10-100 km)
    inertia-gravity waves. Summertime re-stratification weakens submesoscale
    turbulence but enhances inertia-gravity waves near the surface. Thus,
    submesoscale turbulence and inertia-gravity waves undergo vigorous out-of-phase
    seasonal cycles. These results imply a strong seasonal modulation of
    the accuracy of geostrophic velocity diagnosed from submesoscale
    sea-surface height delivered by the Surface Water and Ocean Topography (SWOT)
    satellite mission.
\end{abstract}

@@

% now SWOT
The partition between geostrophic and ageostrophic flows across
the submesoscales and their seasonality have immediate applications for the planning
of the Surface Water and Ocean Topography (SWOT) satellite mission.
SWOT will deliver
sea-surface height (SSH) with about 15 km resolution, thereby providing the
first global SSH measurements at submesoscales \citep{fu_ubelmann2014}.
Thus, the projection of energetic submesoscale ageostrophic flows, such as
inertia-gravity waves (IGWs), onto SSH represents
a challenge --- and an opportunity --- for SWOT. The challenge: how to diagnose
surface velocity from SSH. The opportunity: observations of submesoscale ageostrophic
flows on a global scale.

@@

\begin{figure*}[ht]
  \begin{center}
    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
 \caption{\small LLC4320 wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
          (c) Wavenumber spectrum of SSH variance --- the integral of (a) and (b) over frequency.
          In (a) and (b), the light gray shaded region depicts the dispersion relations for inertia-gravity
          waves from mode 1 through mode 4 across the latitudinal domain; horizontal dashed lines represent the semi-diurnal
          lunar tidal frequency ($M_2$), and the inertial frequency at mid-latitude
          ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). For reference, the gray curvish line represents
          the SWOT baseline requirement $2\times 10^{-4} + 1.25\times 10^{-7} k^{-2}$ m$^2$/cpkm \citep{rodriguez2012}.
          }
  \label{fig4}
 \end{center}
\end{figure*}

Mesoscales flows and the lunar semi-diurnal tide contain most of the
SSH variance (Figures \ref{fig4}a-b). But there is significant SSH variance at
submesoscales, both sub-inertial
and super-inertial. Sub-inertial SSH variance at submesoscales is larger in April, whereas
super-inertial SSH variance at submesoscales is larger in October. At scales smaller
than 100 km (submesoscales), the SSH variance is 66$\%$ larger in October than in April.
IGWs, particularly modes two through four, account for most of the SSH variance
increase in summer --- higher vertical
modes are more sensitive to changes in near-surface stratification (supporting information).
In October, super-inertial IGWs account for a SSH standard deviation of about 4.9 cm,
comparable to the April sub-inertial SSH standard deviation of $\sim$5.2 cm (Table \ref{tab1}).
The
Similarly, the submesoscale super-inertial surface KE is 63$\%$ larger in October
than in April (supporting information).

>>>>

\begin{abstract}
    Recent studies show that the vigorous seasonal cycle of the mixed layer modulates
    upper-ocean submesoscale turbulence.
    Here we provide model-based evidence that the seasonally-changing upper-ocean
    stratification in the Kuroshio Extension also modulates submesoscale  (here 10-100 km)
    inertia-gravity waves. Summertime re-stratification weakens submesoscale
    turbulence but enhances inertia-gravity waves near the surface. Thus,
    submesoscale turbulence and inertia-gravity waves undergo vigorous out-of-phase
    seasonal cycles. These results imply a strong seasonal modulation of
    the accuracy of geostrophic velocity diagnosed from submesoscale
    sea-surface height delivered by the Surface Water and Ocean Topography (SWOT)
    satellite mission.
\end{abstract}

@@

% now SWOT
The partition between geostrophic and ageostrophic flows across
the submesoscales and their seasonality have immediate applications for the planning
of the Surface Water and Ocean Topography (SWOT) satellite mission.
SWOT will deliver
sea-surface height (SSH) with about 15 km resolution, thereby providing the
first global SSH measurements at submesoscales \citep{fu_ubelmann2014}.
Thus, the projection of energetic submesoscale ageostrophic flows, such as
inertia-gravity waves (IGWs), onto SSH represents
a challenge and an opportunity for the SWOT community. The challenge: how to diagnose submesoscale
surface velocity from SSH. The opportunity: stimulation of research on interactions
between geostrophic and ageostrohic flows and their projection onto SSH.

@@

\begin{figure*}[ht]
  \begin{center}
    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
 \caption{\small LLC4320 wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
          (c) Wavenumber spectrum of SSH variance --- the integral of (a) and (b) over frequency.
          In (a) and (b), the light gray shaded region depicts the dispersion relations for inertia-gravity
          waves from mode 1 through mode 4 across the latitudinal domain; horizontal dashed lines represent the semi-diurnal
          lunar tidal frequency ($M_2$), and the inertial frequency at mid-latitude
          ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). For reference, the gray line shows
          the baseline requirement as specified by the SWOT science team \citep{rodriguez2012}.
          }
  \label{fig4}
 \end{center}
\end{figure*}

Mesoscales flows and the lunar semi-diurnal tide contain most of the
SSH variance (Figures \ref{fig4}a-b). But there is significant SSH variance at
submesoscales, both sub-inertial
and super-inertial. Sub-inertial SSH variance at submesoscales is larger in April, whereas
super-inertial SSH variance at submesoscales is larger in October. At scales smaller
than 100 km (submesoscales), the SSH variance is 66$\%$ larger in October than in April.
IGWs, particularly modes two through four, account for most of the SSH variance
increase in summer --- higher vertical
modes are more sensitive to changes in near-surface stratification (supporting information).
In October, super-inertial IGWs account for a SSH standard deviation of about 4.9 cm,
comparable to the April sub-inertial SSH standard deviation of $\sim$5.2 cm (Table \ref{tab1}).
Similarly, the submesoscale super-inertial surface KE is 63$\%$ larger in October
than in April (supporting information).
